---
title: "HARBOR- Urinary Biomarkers"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 2
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
---
```{r,include=FALSE}
knitr::opts_chunk$set(warning=FALSE,message=FALSE)
```



```{r,include=FALSE}
# LOAD PACKAGES
library(rmarkdown)
library(tidyverse)
library(grid)
library(consort)
library(gtsummary)
library(ggpubr)
library(lubridate)
library(purrr)
library(stringr)
library(slider)
library(readr)
library(reshape2)
library(lme4)
library(here)
library(broom)
library(broom.mixed)
library(viridis)
library(nlme)
library(DT)
library(psych)
library(knitr)
library(networkD3)
library(htmlwidgets)
library(pBrackets)
library(gt)
library(ggcorrplot)
library(car)
```

```{r,include=FALSE}
# LOAD DATA 

#covariates=read.csv(here::here("data","STSvariablesDEID.7.22.24.csv"))
covariates=read.csv(here::here("data","STSvariablesDEID.7.20.25.csv"))
ortimes= read.csv(here::here("data","ORtimesDEID.csv"))
#biomarkers= read.csv(here::here("data","All_BiomarkerLongformHARBORAKI.9.1.24.csv"))

biomarkers= read.csv(here::here("data","Allbiomarkers.ONLY.longform.1.16.25.V3.FINAL.csv"))
hemodynamics= read_rds(here::here("data", "hemodynamics.rds"))

akidata= read.csv(here::here("data", "HARBORAKI.128patients.AKI48.DO2.12.13.42.csv"))

timestamp= read.csv("/scratch4/nfarada1/Marina/HARBOR-UrineBiomarkers/data/1.16.25.biomarker timepoints.DEID.csv")

crystalloids= read.csv("/scratch4/nfarada1/Marina/HARBOR-UrineBiomarkers/data/DEID_totalcrystalloids_upto12.31.25.csv")%>% slice(-2201) #remove wrong/duplicate row

vasopressors = read.csv("/scratch4/nfarada1/Marina/HARBOR-UrineBiomarkers/data/vasopressor-inotrope-dose.csv")

```


```{r,include=FALSE}
# PROCESS DATA


datasets <- list(covariates, hemodynamics, ortimes, akidata, biomarkers)
lapply(datasets, function(df) df$harbor_id <- as.factor(df$harbor_id))
biomarkers$timepoint <- as.factor(biomarkers$timepoint)

covariates= covariates[1:128, ]
covariates <- covariates %>%
  left_join(ortimes, by = "harbor_id") %>%
  left_join(akidata %>% select("harbor_id","AKI_48","X48hrs_delCr"), by = "harbor_id")


covariates= covariates %>% mutate(rfchroniclungdisease= ifelse(rfchroniclungdisease=="No","No","Yes"), 
                                  medsbetablockerswithin24hours= ifelse(medsbetablockerswithin24hours=="Yes","Yes","No"),
                                  rfpriorcva=ifelse(rfpriorcva=="Yes","Yes","No"))



covariates <- covariates %>%
  mutate(
    packed_rbc = case_when(
      is.na(intraopbloodproductsrbcunits)     ~ "0",           # NA becomes "0"
      intraopbloodproductsrbcunits > 1        ~ ">1",          # Values > 1 become ">1"
      TRUE         ~ as.character(intraopbloodproductsrbcunits)  # Keep 0, 1 as-is
    ),
    packed_rbc = factor(packed_rbc, levels = c("0", "1", ">1"))  # set factor levels
  )





##########

# change picograms(pg) to nanograms (ng)
biomarkers$NGAL=biomarkers$NGAL_pgmL/1000
biomarkers$YKL=biomarkers$YKL40_pgmL/1000

# divide biomarkers by UCR
biomarkers$NGAL_UCR=biomarkers$NGAL/biomarkers$URCREA_mgdL # note the difference in units
biomarkers$YKL_UCR=biomarkers$YKL/biomarkers$URCREA_mgdL # note the difference in units
biomarkers$AL_UCR=biomarkers$mAlb_mgdL/biomarkers$URCREA_mgdL
biomarkers$NGAL_YKL= biomarkers$NGAL/biomarkers$YKL
biomarkers$NGAL_YKL_UCR=biomarkers$NGAL_YKL/biomarkers$URCREA_mgdL


# log 2 biomarkers
biomarkers$log2NGAL=log(biomarkers$NGAL, base = 2)
biomarkers$log2NGAL_UCR=log(biomarkers$NGAL_UCR, base = 2)
biomarkers$log2YKL=log(biomarkers$YKL, base = 2)
biomarkers$log2YKL_UCR=log(biomarkers$YKL_UCR, base = 2)
biomarkers$log2AL_UCR=log(biomarkers$AL_UCR, base = 2)
biomarkers$log2_NGAL_YKL= log(biomarkers$NGAL_YKL, base=2)
biomarkers$log2_NGAL_YKL_UCR= log(biomarkers$NGAL_YKL_UCR, base=2)

biomarkers$log2akiscore=log(as.numeric(ifelse(biomarkers$akiscore=="< 0.04", 0.02, biomarkers$akiscore),base=2))
biomarkers$akiscore=ifelse(biomarkers$akiscore=="< 0.04", 0, biomarkers$akiscore)
biomarkers$akiscore=as.numeric(biomarkers$akiscore)
  
# Delta biomarkers 

biowide=biomarkers %>% select(timepoint, harbor_id,NGAL, YKL, NGAL_UCR, YKL_UCR, AL_UCR, log2NGAL, log2NGAL_UCR,log2YKL_UCR,log2AL_UCR,akiscore,log2akiscore)
biowide=biowide %>% pivot_wider(names_from = timepoint,  values_from = c(NGAL,YKL, NGAL_UCR, YKL_UCR, AL_UCR,log2NGAL, log2NGAL_UCR,log2YKL_UCR,log2AL_UCR,akiscore,log2akiscore))

biowide= biowide %>% mutate(NGAL_2_0= NGAL_2-NGAL_0,
                            log2NGAL_2_0= log2NGAL_2- log2NGAL_0,
                            NGAL_UCR_2_0 = NGAL_UCR_2-NGAL_UCR_0,
                            log2NGAL_UCR_0.5_0 = log2NGAL_UCR_0.5-log2NGAL_UCR_0,
                            log2NGAL_UCR_2_0 = log2NGAL_UCR_2-log2NGAL_UCR_0,
                            log2NGAL_2_1= log2NGAL_2- log2NGAL_1,
                            log2NGAL_UCR_2_1 = log2NGAL_UCR_2-log2NGAL_UCR_1,
                            log2NGAL_UCR_3_0 = log2NGAL_UCR_3-log2NGAL_UCR_0,
                            log2NGAL_UCR_1_0.5 = log2NGAL_UCR_1-log2NGAL_UCR_0.5,
                            log2NGAL_UCR_3_2 = log2NGAL_UCR_3-log2NGAL_UCR_2,
                            NGAL_1_0= NGAL_1-NGAL_0,
                            log2NGAL_1_0= log2NGAL_1-log2NGAL_0,
                            NGAL_UCR_1_0 = NGAL_UCR_1-NGAL_UCR_0,
                            log2NGAL_UCR_1_0 = log2NGAL_UCR_1-log2NGAL_UCR_0,
                            YKL_UCR_2_0 = YKL_UCR_2-YKL_UCR_0,
                            log2YKL_UCR_2_0 = log2YKL_UCR_2-log2YKL_UCR_0,
                            YKL_UCR_1_0 = YKL_UCR_1-YKL_UCR_0,
                            log2YKL_UCR_0.5_0 = log2YKL_UCR_0.5-log2YKL_UCR_0,
                            log2YKL_UCR_1_0 = log2YKL_UCR_1-log2YKL_UCR_0,
                            log2YKL_UCR_2_1 = log2YKL_UCR_2-log2YKL_UCR_1,
                            log2YKL_UCR_3_0 = log2YKL_UCR_3-log2YKL_UCR_0,
                            log2YKL_UCR_1_0.5 = log2YKL_UCR_1-log2YKL_UCR_0.5,
                            log2YKL_UCR_3_2 = log2YKL_UCR_3-log2YKL_UCR_2,
                            AL_UCR_2_0 = AL_UCR_2-AL_UCR_0,
                            log2AL_UCR_2_0 = log2AL_UCR_2-log2AL_UCR_0,
                            AL_UCR_1_0 = AL_UCR_1-AL_UCR_0,
                            log2AL_UCR_1_0 = log2AL_UCR_1-log2AL_UCR_0,
                            akiscore_1_0= akiscore_1-akiscore_0,
                            akiscore_2_0= akiscore_2-akiscore_0,
                            akiscore_3_0= akiscore_3-akiscore_0,
                            log2akiscore_1_0= log2akiscore_1-log2akiscore_0,
                            log2akiscore_2_0= log2akiscore_2-log2akiscore_0,
                            log2akiscore_3_0= log2akiscore_3-log2akiscore_0,
                            log2akiscore_0.5_0= log2akiscore_0.5-log2akiscore_0,
                            log2akiscore_1_0.5= log2akiscore_1-log2akiscore_0.5,
                            log2akiscore_2_1= log2akiscore_2-log2akiscore_1,
                            log2akiscore_3_2= log2akiscore_3-log2akiscore_2
                            )





# calculate proctime
start_time_posix <- as.POSIXct(ortimes$anes_start_time, format = "%m/%d/%Y %H:%M", tz = "UTC")
end_time_posix <- as.POSIXct(ortimes$anes_end_time, format = "%m/%d/%Y %H:%M", tz = "UTC")
time_diff <- as.numeric(end_time_posix - start_time_posix,units="mins")
covariates$proctime=time_diff


#CPB time using cpb1on and last cpb time

df_clean <- hemodynamics %>%
  select(harbor_id, cpb1on, cpboff_last) %>% 
  distinct()

start_time_posix <- as.POSIXct(df_clean$cpb1on, format = "%m/%d/%Y %H:%M", tz = "UTC")
end_time_posix <- as.POSIXct(df_clean$cpboff_last, format = "%m/%d/%Y %H:%M", tz = "UTC")
time_diff <- as.numeric(end_time_posix - start_time_posix,units="mins")
covariates$cpbtime=time_diff


hemodynamics = hemodynamics %>% mutate(state= case_when(CVP>12 ~ "R",
                                                         MAP<=65 & CVP>8 ~ "R",
                                                         MAP<=55 & !is.na(CVP)~ "R",
                                                         MAP>75 & CVP<=8 ~"B",
                                                         MAP>85 & CVP<=10 ~"B",
                                                         MAP>65 & CVP>10 & CVP<=12 ~"W",
                                                         MAP>55 & MAP<=75 & CVP<=8 ~"W",
                                                         MAP>65 & MAP<=85 & CVP>8 & CVP<=10~"W"))

hemodynamics = hemodynamics %>% mutate(zone5= case_when(MAP>45 & MAP<=55 & CVP >8 ~"zone5",
                                                        MAP>55 & MAP<=65 & CVP >12 ~"zone5",
                                                        MAP>65 & MAP<=75 & CVP >16 ~"zone5",
                                                        MAP>75 & MAP<=85 & CVP >16 ~"zone5",
                                                        MAP>85 & MAP<=95 & CVP >18 ~"zone5",
                                                        MAP>95 & MAP<=105 & CVP >18 ~"zone5",
                                                        MAP>105 & MAP<=115 & CVP >18 ~"zone5"))


# Summarize the data
datasum <- hemodynamics %>%
  mutate(
    lowci = CI < 2,
    lowmap = MAP < 65,
    lowcimap= CI<2 & MAP<65,
    highcvp=CVP>12,
    lowmap55 = MAP <55,
    lowcihighcvp= CI<2 & CVP>12
  )%>%
  group_by(harbor_id) %>%
  summarise(
    lowcimins = sum(lowci, na.rm = TRUE) / 12,
    lowmap55mins = sum(lowmap55, na.rm = TRUE) / 12,
    lowcimapmins= sum(lowcimap, na.rm = TRUE) / 12,
    lowcihighcvpmins= sum(lowcihighcvp,na.rm=TRUE)/12,
    prelowcimins = sum(lowci * (wftime < cpb1on), na.rm = TRUE) / 12,
    postlowcimins = sum(lowci * (wftime > cpboff_last), na.rm = TRUE) / 12,
    prelowmap55mins = sum(lowmap55 * (wftime < cpb1on), na.rm = TRUE) / 12,
    postlowmap55mins = sum(lowmap55 * (wftime > cpboff_last), na.rm = TRUE) / 12,
    prelowcimapmins = sum(lowcimap * (wftime < cpb1on), na.rm = TRUE) / 12,
    postlowcimapmins = sum(lowcimap * (wftime > cpboff_last), na.rm = TRUE) / 12,
    prelowcihighcvpmins = sum(lowcihighcvp * (wftime < cpb1on), na.rm = TRUE) / 12,
    postlowcihighcvpmins = sum(lowcihighcvp * (wftime > cpboff_last), na.rm = TRUE) / 12, 
    highcvpmins= sum(highcvp,na.rm = TRUE)  / 12,
    prehighcvpmins= sum((highcvp) * (wftime < cpb1on), na.rm = TRUE) / 12,
    posthighcvpmins= sum((highcvp) * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrahighcvpmins= sum((highcvp) * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_highcvpmins= sum((highcvp) * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    lowmapmins= sum(lowmap,na.rm = TRUE)  / 12,
    prelowmapmins= sum((lowmap) * (wftime < cpb1on), na.rm = TRUE) / 12,
    postlowmapmins= sum((lowmap) * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intralowmapmins= sum((lowmap) * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_lowmapmins= sum((lowmap) * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    redmins= sum(state=="R", na.rm = TRUE)  / 12,
    red5mins= sum(state=="R", na.rm = TRUE)  / 60,
    preredmins= sum((state=="R") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postredmins= sum((state=="R") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intraredmins= sum((state=="R") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_redmins= sum((state=="R") * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    preintra_red5mins= sum((state=="R") * (wftime <= cpboff_last), na.rm = TRUE) / 60,
    bluemins= sum(state=="B", na.rm = TRUE)  / 12,
    prebluemins= sum((state=="B") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postbluemins= sum((state=="B") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrabluemins= sum((state=="B") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_bluemins= sum((state=="B") * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    whitemins= sum(state=="W", na.rm = TRUE)  / 12,
    prewhitemins= sum((state=="W") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postwhitemins= sum((state=="W") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrawhitemins= sum((state=="W") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_whitemins= sum((state=="W") * (wftime <= cpboff_last), na.rm = TRUE) / 12, zone5mins= sum(zone5=="zone5", na.rm = TRUE)  / 12,
    prezone5mins= sum((zone5=="zone5") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postzone5mins= sum((zone5=="zone5") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrazone5mins= sum((zone5=="zone5") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    white5mins=sum(state=="W", na.rm = TRUE)  / 60,
    blue5mins=sum(state=="B", na.rm = TRUE)  / 60,
    MAPless65_5mins=sum(MAP < 65, na.rm = TRUE)/60, 
    MAPge65_5mins=sum(MAP >= 65, na.rm = TRUE)/60,
    MAPless55_5mins=sum(MAP <55, na.rm = TRUE)/60,
    MAPge55_64_5mins=sum(MAP >= 55 & MAP<64, na.rm = TRUE)/60,
    
    blue5mins_nobypass=(prebluemins+postbluemins)/5,
    white5mins_nobypass=(prewhitemins+postwhitemins)/5,
    red5mins_nobypass=(preredmins+postredmins)/5,
    intrablue5mins= intrabluemins/5,
    intrawhite5mins=intrawhitemins/5,
    intrared5mins=intraredmins/5
    
    ) %>%
  arrange(harbor_id)

biowide=datasum %>% left_join(biowide, by="harbor_id")

biomarkers <- biomarkers %>%
  mutate(
    # 1. Force timepoint to be a number
    timepoint = as.numeric(as.character(timepoint)),
    # 2. Match the numbers to labels
    timepoint = case_match(timepoint, 
                           0   ~ "Baseline", 
                           0.5 ~ "PreCPB", 
                           1   ~ "OffCPB", 
                           2   ~ "ICU", 
                           3   ~ "24hICU")
  )

biowide= biowide %>% left_join(covariates,by="harbor_id")

biowide$AKI_48=as.factor(ifelse(biowide$AKI_48=="YES",1,0))

biomarkers=biomarkers %>% left_join(biowide%>%select(harbor_id,AKI_48,cpbtime,proctime),by="harbor_id")




timestamp$anesthesia.start.time <- strptime(timestamp$anesthesia.start.time, format = "%H:%M")
timestamp$Timestamp.0_baseline <- strptime(timestamp$Timestamp.0_baseline, format = "%H:%M")
timestamp$Timestamp.1_offCPB<- strptime(timestamp$Timestamp.1_offCPB, format = "%H:%M")
timestamp$Timestamp.2_ICUarrival <- strptime(timestamp$Timestamp.2_ICUarrival, format = "%H:%M")
timestamp$Timestamp.3_ICUnextday <- as.POSIXct(strptime(timestamp$Timestamp.3_ICUnextday, format = "%H:%M"))+86400



timestamp$Timestamp.0_baseline <- as.numeric(difftime(timestamp$Timestamp.0_baseline, timestamp$Timestamp.1_offCPB, units = "mins"))
timestamp$Timestamp.2_ICUarrival <- as.numeric(difftime(timestamp$Timestamp.2_ICUarrival, timestamp$Timestamp.1_offCPB, units = "mins"))
timestamp$Timestamp.3_ICUnextday <- as.numeric(difftime(timestamp$Timestamp.3_ICUnextday, timestamp$Timestamp.1_offCPB, units = "mins"))
timestamp$Timestamp.1_offCPB <- as.numeric(difftime(timestamp$Timestamp.1_offCPB, timestamp$Timestamp.1_offCPB, units = "mins"))


colnames(timestamp)=c("harbor_id","anes_start","Baseline","OffCPB","ICU","24hICU")

biowide$log2NGAL_UCR_2_1_avg= (biowide$log2NGAL_UCR_2+biowide$log2NGAL_UCR_1)/2
biowide$log2YKL_UCR_2_1_avg=(biowide$log2NGAL_UCR_2+biowide$log2NGAL_UCR_1)/2

rm(akidata,datasets,ortimes,df_clean,datasum)

hemodynamics=hemodynamics %>% mutate(cat_cpb=ifelse(wftime<cpb1on,"pre",ifelse(wftime>cpboff_last,"post","intra")))
hemodynamics$cat_cpb= factor(hemodynamics$cat_cpb,levels=c("pre","intra","post"))

hemodynamics= hemodynamics %>% left_join(biowide %>% select(AKI_48,X48hrs_delCr,redmins, harbor_id,proctime),by="harbor_id")

biowide = biowide %>% left_join(crystalloids, by="sickbayid")

biowide = biowide %>% left_join(vasopressors %>% rename(sickbayid=ID,vasopressor=total ) %>% select(sickbayid, vasopressor), by="sickbayid") %>% mutate(vasopressor=vasopressor/1000)
```




```{r}
# Calculate proportion of missingness for variables 'x' and 'y' per harbor_id
missingness_summary <- hemodynamics %>%
  group_by(harbor_id) %>%
  summarise(
    # mean(is.na(...)) calculates the proportion of TRUE (missing) values
    prop_missing_x = mean(is.na(MAP)),
    prop_missing_y = mean(is.na(CVP)),
    
    # Optional: Calculate total row count per ID to keep context of the proportions
    total_records = n()
  )

```



# Old Biomarker Analyses

## Patient exclusion 

```{r,echo=FALSE}
# PATIENT EXCLUSION 

ids= covariates$harbor_id
# Subset data 
remove_ids=c("3","86","10","32","50","60","69")
biomarkers= biomarkers %>% filter(!harbor_id %in% remove_ids) %>% filter(harbor_id %in% ids)
biowide= biowide %>% filter(!harbor_id %in% remove_ids)%>% filter(harbor_id %in% ids)
hemodynamics= hemodynamics %>% filter(!harbor_id %in% remove_ids)%>% filter(harbor_id %in% ids)
rm(covariates,end_time_posix,g,ids,remove_ids,start_time_posix,time_diff)




# Exclusion diagram 

g=add_box(txt="HARBOR-AKI Population (n=128)")|>
  add_side_box(txt="Excluded (n=2):\n\u2022 >1 Aortic cross-clamp")|>
  add_box(txt="Patients (n=126)")|>
  add_side_box(txt="Excluded (n=4):\n\u2022 Missing urinary biomarkers")|>
  add_box(txt="Patients (n=122)")|>
  add_side_box(txt="Excluded (n=1):\n\u2022 Intra-Aortic Balloon Pump")|>
  add_box(txt="Patients (n=121)")
plot(g)
```


```{r}
# Calculate proportion of missingness for variables 'x' and 'y' per harbor_id
missingness_summary <- hemodynamics %>%
  group_by(harbor_id) %>%
  summarise(
    # mean(is.na(...)) calculates the proportion of TRUE (missing) values
    prop_missing_x = mean(is.na(MAP)),
    prop_missing_y = mean(is.na(CVP)),
    
    # Optional: Calculate total row count per ID to keep context of the proportions
    total_records = n()
  )

```




## Covariates/ Zone Exposures 



Total Surgery Minutes in Red Zone, Tertiles

```{r,echo=FALSE}


biowide <- biowide %>%
  mutate(akiscore_3_thresh03 = factor(
    if_else(akiscore_3  > 0.3, ">0.3", "≤0.3"),
    levels = c("≤0.3", ">0.3")
  ), akiscore_3_thresh2 = factor(
    if_else(akiscore_3  > 2, ">2", "≤2"),
    levels = c("≤2", ">2")
  ))

biowide <- biowide %>%
  mutate(akiscore_2_thresh03 = factor(
    if_else(akiscore_2  > 0.3, ">0.3", "≤0.3"),
    levels = c("≤0.3", ">0.3")
  ), akiscore_2_thresh2 = factor(
    if_else(akiscore_2  > 2, ">2", "≤2"),
    levels = c("≤2", ">2")
  ))



quantile(biowide$redmins, probs = c(0, 1/3, 2/3, 1))

biowide$redzonetertile <- cut(biowide$redmins,
                  breaks = quantile(biowide$redmins, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
                  include.lowest = TRUE,
                  labels = c("<77", "77-128", ">128"))

covs=biowide %>% dplyr::select(redzonetertile,patientage,gender,bmi,rfhypertension,rfdiabetes,priormi,hemodataef,rfpriorcva ,rfchroniclungdisease,medsbetablockerswithin24hours, rfhemoglobin,rflastcreatlevel,proctime,cpbtime,heartfailure,akiscore_2_thresh03,akiscore_2_thresh2,akiscore_3_thresh03,akiscore_3_thresh2, log2NGAL_UCR_2, NGAL_UCR_2, akiscore_2, akiscore_3)

covs %>% tbl_summary(by=redzonetertile,statistic=all_continuous()~"{mean} ({sd})", missing_text = "Missing")

```

```{r}

quantile(biowide$bluemins, probs = c(0, 1/3, 2/3, 1))

biowide$bluezonetertile <- cut(biowide$bluemins,
                  breaks = quantile(biowide$bluemins, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
                  include.lowest = TRUE,
                  labels = c("<24", "24-61", ">61"))

covs=biowide %>% dplyr::select(bluezonetertile,patientage,gender,bmi,rfhypertension,rfdiabetes,priormi,hemodataef,rfpriorcva ,rfchroniclungdisease,medsbetablockerswithin24hours, rfhemoglobin,rflastcreatlevel,proctime,cpbtime,heartfailure,akiscore_2_thresh03,akiscore_2_thresh2,akiscore_3_thresh03,akiscore_3_thresh2, log2NGAL_UCR_2, NGAL_UCR_2, akiscore_2, akiscore_3)

covs %>% tbl_summary(by=bluezonetertile,statistic=all_continuous()~"{mean} ({sd})", missing_text = "Missing")

```




## Change in biomarkers during surgery {.tabset}



```{r,echo=FALSE}
biomarkers %>%
  drop_na(AKI_48) %>%
  ggplot(aes(x = timepoint, y = log2NGAL_UCR)) +
  geom_boxplot() +
  stat_summary(fun.data=function(x) data.frame(y=8,label=paste0("n=",length(x))),geom="text",vjust=-0.5)+ylab("log2(NGAL/UCR)")+xlab("Timepoint")
```



Paired t-test between baseline and other timepoints:

```{r,echo=FALSE}
comparisons <- list(
  PreCPB  = biowide$log2NGAL_UCR_0.5,
  OffCPB  = biowide$log2NGAL_UCR_1,
  ICU     = biowide$log2NGAL_UCR_2,
  `24hICU` = biowide$log2NGAL_UCR_3
)

t_test_results <- imap_dfr(comparisons, function(x, name) {
  
  test <- t.test(x,biowide$log2NGAL_UCR_0,  paired = TRUE)
  
  tibble(
    timepoint = name,
    mean_difference = unname(test$estimate),   # mean of differences
    conf_low = test$conf.int[1],
    conf_high = test$conf.int[2],
    t_statistic = unname(test$statistic),
    p_value = test$p.value
  )
})

# Round for reporting
t_test_results %>%
  mutate(across(where(is.numeric), ~ round(., 4)))
```


```{r,echo=FALSE}
comparisons <- list(
  PreCPB  = biowide$log2NGAL_UCR_0.5,
  OffCPB  = biowide$log2NGAL_UCR_1,
  ICU     = biowide$log2NGAL_UCR_2,
  `24hICU` = biowide$log2NGAL_UCR_3
)
t_test_results <- imap_dfr(comparisons, function(x, name) {
  
  # Swap the order here: (Post-treatment, Baseline)
  # This makes the result: (Mean of x) - (Mean of Base)
  test <- t.test(x, biowide$log2NGAL_UCR_0, paired = FALSE)
  
  tibble(
    timepoint = name,
    # Now estimate[1] is 'x' and estimate[2] is 'Base'
    # This subtraction gives: Post - Base
    mean_difference = test$estimate[1] - test$estimate[2],
    conf_low = test$conf.int[1],
    conf_high = test$conf.int[2],
    t_statistic = unname(test$statistic),
    p_value = test$p.value
  )
})

# Display results
t_test_results %>%
  mutate(across(where(is.numeric), ~ round(., 4)))
```

```{r,echo=FALSE}
# OffCPB and ICU
test <- t.test(biowide$log2NGAL_UCR_2, biowide$log2NGAL_UCR_1, paired = FALSE)
  
  tibble(
    mean_difference = test$estimate[1] - test$estimate[2],
    conf_low = test$conf.int[1],
    conf_high = test$conf.int[2],
    t_statistic = unname(test$statistic),
    p_value = test$p.value
  )
  
  
  test <- t.test(biowide$log2NGAL_UCR_2, biowide$log2NGAL_UCR_1,  paired = TRUE)
  
  tibble(
    mean_difference = unname(test$estimate),   # mean of differences
    conf_low = test$conf.int[1],
    conf_high = test$conf.int[2],
    t_statistic = unname(test$statistic),
    p_value = test$p.value
  )
```




```{r,echo=FALSE}
test <- t.test(biowide$log2NGAL_UCR_3, biowide$log2NGAL_UCR_2, paired = FALSE)
  
  tibble(
    mean_difference = test$estimate[1] - test$estimate[2],
    conf_low = test$conf.int[1],
    conf_high = test$conf.int[2],
    t_statistic = unname(test$statistic),
    p_value = test$p.value
  )
  
  
  test <- t.test(biowide$log2NGAL_UCR_3, biowide$log2NGAL_UCR_2,  paired = TRUE)
  
  tibble(
    mean_difference = unname(test$estimate),   # mean of differences
    conf_low = test$conf.int[1],
    conf_high = test$conf.int[2],
    t_statistic = unname(test$statistic),
    p_value = test$p.value
  )
```












```{r,echo=FALSE}
test <- t.test(biowide$akiscore_3,biowide$akiscore_0,  paired = TRUE)
  
  tibble(
    timepoint = "24hICU",
    mean_difference = unname(test$estimate),   # mean of differences
    conf_low = test$conf.int[1],
    conf_high = test$conf.int[2],
    t_statistic = unname(test$statistic),
    p_value = test$p.value
  )
  
  
  
  
  test <- t.test(biowide$akiscore_3,biowide$akiscore_0,  paired = FALSE)
  
  tibble(
    timepoint = "24hICU",
    # Now estimate[1] is 'x' and estimate[2] is 'Base'
    # This subtraction gives: Post - Base
    mean_difference = test$estimate[1] - test$estimate[2],
    conf_low = test$conf.int[1],
    conf_high = test$conf.int[2],
    t_statistic = unname(test$statistic),
    p_value = test$p.value
  )
```


```{r,echo=FALSE}
mw_test <- wilcox.test(biowide$akiscore_3, biowide$akiscore_0, 
                       paired = FALSE, 
                       conf.int = TRUE)

# Cleanly report the results
mw_test %>%
  tidy() %>%
  select(
    statistic, 
    p_value = p.value, 
    estimate_median_diff = estimate, # This is the Hodges-Lehmann estimator
    conf_low = conf.low, 
    conf_high = conf.high
  ) %>%
  mutate(across(where(is.numeric), ~ round(., 4)))


mw_test <- wilcox.test(biowide$akiscore_3, biowide$akiscore_0, 
                       paired = TRUE, 
                       conf.int = TRUE)

# Cleanly report the results
mw_test %>%
  tidy() %>%
  select(
    statistic, 
    p_value = p.value, 
    estimate_median_diff = estimate, # This is the Hodges-Lehmann estimator
    conf_low = conf.low, 
    conf_high = conf.high
  ) %>%
  mutate(across(where(is.numeric), ~ round(., 4)))
```


Paired t-test between baseline and other timepoints YKL:
```{r,echo=FALSE}
t_test_results <- tibble(
  timepoint = c("PreCPB", "OffCPB", "ICU", "24hICU"),
  p_value = c(
    t.test(biowide$log2YKL_UCR_0, biowide$log2YKL_UCR_0.5, paired = TRUE)$p.value,
    t.test(biowide$log2YKL_UCR_0, biowide$log2YKL_UCR_1, paired = TRUE)$p.value,
    t.test(biowide$log2YKL_UCR_0, biowide$log2YKL_UCR_2, paired = TRUE)$p.value,
    t.test(biowide$log2YKL_UCR_0, biowide$log2YKL_UCR_3, paired = TRUE)$p.value
  ),
  t_statistic = c(
    t.test(biowide$log2YKL_UCR_0, biowide$log2YKL_UCR_0.5, paired = TRUE)$statistic,
    t.test(biowide$log2YKL_UCR_0, biowide$log2YKL_UCR_1, paired = TRUE)$statistic,
    t.test(biowide$log2YKL_UCR_0, biowide$log2YKL_UCR_2, paired = TRUE)$statistic,
    t.test(biowide$log2YKL_UCR_0, biowide$log2YKL_UCR_3, paired = TRUE)$statistic
  )
)

print(t_test_results %>% mutate(
    p_value = round(p_value, 4),
    t_statistic = round(t_statistic, 4)))
```




Paired t-test between baseline and other timepoints Nephrocheck:
```{r,echo=FALSE}
t_test_results <- tibble(
  timepoint = c("OffCPB", "ICU", "24hICU"),
  p_value = c(
    t.test(biowide$log2akiscore_0, biowide$log2akiscore_1, paired = TRUE)$p.value,
    t.test(biowide$log2akiscore_0, biowide$log2akiscore_2, paired = TRUE)$p.value,
    t.test(biowide$log2akiscore_0, biowide$log2akiscore_3, paired = TRUE)$p.value
  ),
  t_statistic = c(
    t.test(biowide$log2akiscore_0, biowide$log2akiscore_1, paired = TRUE)$statistic,
    t.test(biowide$log2akiscore_0, biowide$log2akiscore_2, paired = TRUE)$statistic,
    t.test(biowide$log2akiscore_0, biowide$log2akiscore_3, paired = TRUE)$statistic
  )
)

print(t_test_results %>% mutate(
    p_value = round(p_value, 4),
    t_statistic = round(t_statistic, 4)))

t.test(biowide$akiscore_1, biowide$akiscore_0, paired = TRUE)
t.test(biowide$akiscore_2, biowide$akiscore_0, paired = TRUE)
t.test(biowide$akiscore_3, biowide$akiscore_0, paired = TRUE)
```














## Line/Boxplots biomarkers raw and log over timepoints


```{r,echo=FALSE}
biomarkers %>% ggplot(aes(x=timepoint,y=NGAL_UCR))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    labs(x = "Timepoint", y = "NGAL/UCR") 

biomarkers %>% ggplot(aes(x=timepoint,y=NGAL_UCR))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    geom_boxplot(alpha = 1, width = 0.4, outlier.size = 1.5,fill="white")+
    labs(x = "Timepoint", y = "NGAL/UCR") 

biomarkers %>% ggplot(aes(x=timepoint,y=log2NGAL_UCR))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    geom_boxplot(alpha = 1, width = 0.4, outlier.size = 1.5,fill="white")+
    labs(x = "Timepoint", y = "log2(NGAL/UCR)") 

biomarkers %>% ggplot(aes(x=timepoint,y=log2NGAL_UCR))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    geom_boxplot(alpha = 0.5, width = 0.4, outlier.size = 1.5)+
    labs(x = "Timepoint", y = "log2(NGAL/UCR)") 





biomarkers %>% ggplot(aes(x=timepoint,y=YKL_UCR))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    labs(x = "Timepoint", y = "YKL/UCR") 

biomarkers %>% ggplot(aes(x=timepoint,y=YKL_UCR))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    geom_boxplot(alpha = 1, width = 0.4, outlier.size = 1.5,fill="white")+
    labs(x = "Timepoint", y = "YKL/UCR") 

biomarkers %>% ggplot(aes(x=timepoint,y=log2YKL_UCR))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    geom_boxplot(alpha = 1, width = 0.4, outlier.size = 1.5,fill="white")+
    labs(x = "Timepoint", y = "log2(YKL/UCR)") 

biomarkers %>% ggplot(aes(x=timepoint,y=log2YKL_UCR))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    geom_boxplot(alpha = 0.5, width = 0.4, outlier.size = 1.5)+
    labs(x = "Timepoint", y = "log2(YKL/UCR)") 







biomarkers %>% ggplot(aes(x=timepoint,y=akiscore))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    labs(x = "Timepoint", y = "Nephrocheck") 

biomarkers %>% ggplot(aes(x=timepoint,y=akiscore))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    geom_boxplot(alpha = 1, width = 0.4, outlier.size = 1.5,fill="white")+
    labs(x = "Timepoint", y = "Nephrocheck") 

biomarkers %>% ggplot(aes(x=timepoint,y=log2akiscore))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    geom_boxplot(alpha = 1, width = 0.4, outlier.size = 1.5,fill="white")+
    labs(x = "Timepoint", y = "log2(Nephrocheck)") 

biomarkers %>% ggplot(aes(x=timepoint,y=log2akiscore))+
    geom_line(aes(group=harbor_id),alpha = 0.3, color = "black") + 
    geom_boxplot(alpha = 0.5, width = 0.4, outlier.size = 1.5)+
    labs(x = "Timepoint", y = "log2(Nephrocheck)") 
```





## Swihart Lasagna Plots {.tabset}



### Lasagna
```{r,echo=FALSE}
hemodynamics <- hemodynamics %>%
  group_by(harbor_id) %>%
  mutate(time_seconds = (row_number()-1) * 5) %>%
  ungroup() %>%
  mutate(mins = time_seconds / 60)


hemodynamics <- hemodynamics %>%
  mutate(harbor_id2 = factor(harbor_id)) %>% droplevels()


hemodynamics$state <- ifelse(is.na(hemodynamics$state), "Missing", hemodynamics$state)
hemodynamics$state <- factor(hemodynamics$state, levels = c("B", "W", "R", "Missing"))
```


```{r,echo=FALSE}
plot=hemodynamics %>%ggplot(aes(x = mins, y = harbor_id2, fill = state))+geom_tile()+xlab("Minutes from Start of Surgery")+theme_minimal()+theme(axis.text.y=element_blank(),axis.ticks.y=element_blank(),axis.line.y=element_blank(),panel.grid.major.y=element_blank(),panel.grid.minor.y=element_blank())+ylab("Patient")+scale_fill_manual(
    values = c("R" = "#DC3912", "W" = "lavender", "B" = "#3366CC","Missing" = "gray50"),
    labels = c("R" = "Zone 3 - High Risk", "W" = "Zone 2 - Intermediate Risk", "B" = "Zone 1 - Low Risk", "Missing"="Missing"),
    name = "Zones"  
  )
plot

ggsave("zones.png", 
       plot = plot, 
       device = "png", 
       width = 10, 
       height = 6, 
       units = "in", 
       dpi = 400)

ggsave("zones.jpg", 
       plot = plot, 
       device = "jpeg", 
       width = 10, 
       height = 6, 
       units = "in", 
       dpi = 400)
```



# New Biomarker Analyses

## Red/White/Blue Zones {.tabset}
### AKI

```{r,echo=FALSE}
tidy(glm(AKI_48 ~ red5mins+white5mins+blue5mins+predmort, data = biowide, family = "binomial"))
```



```{r, include=FALSE, evaluate=FALSE}
m1 <- glm(AKI_48 ~ red5mins + white5mins + blue5mins + predmort, data = biowide, family = "binomial")
vif(m1)
```




```{r,echo=FALSE}

tidy(glm(AKI_48 ~ red5mins+white5mins+blue5mins+predmort+MAPless65_5mins, data = biowide, family = "binomial"))
```






```{r, include=FALSE, evaluate=FALSE}
m1 <- glm(AKI_48 ~ red5mins+white5mins+blue5mins+predmort+MAPless65_5mins, data = biowide, family = "binomial")
vif(m1)
```


```{r,echo=FALSE}
tidy(glm(AKI_48 ~ red5mins+white5mins+blue5mins+predmort+cpbtime, data = biowide, family = "binomial"))
```




```{r, include=FALSE, evaluate=FALSE}
m1 <- glm(AKI_48 ~ red5mins+white5mins+blue5mins+predmort+cpbtime, data = biowide, family = "binomial")
vif(m1)
```



### Mixed effects  NGAL/cr through ICU arrival

```{r,echo=FALSE}
biomarkers2= biomarkers %>% filter(!timepoint  %in% c("PreCPB","24hICU")) %>% select(harbor_id,timepoint,log2NGAL_UCR)
timepoint_1_values <- biomarkers2 %>%
  filter(timepoint == "Baseline") %>%
  select(harbor_id, baseline_value = log2NGAL_UCR)
biomarkers2 = biomarkers2%>%
  filter(timepoint != "Baseline") %>%
  left_join(timepoint_1_values, by = "harbor_id")

offcpb_hemo = hemodynamics %>% group_by(harbor_id) %>% filter(wftime<=cpboff_last) %>% mutate(harbor_id=as.integer(harbor_id))%>% summarise(redzone5mins=sum(state=="R", na.rm = TRUE)  / 60,whitezone5mins=sum(state=="W", na.rm = TRUE)  / 60,bluezone5mins=sum(state=="B", na.rm = TRUE)  / 60, MAPless65_5mins= sum(MAP<65,na.rm=TRUE)/60, MAPge65_5mins=sum(MAP >= 65, na.rm = TRUE)/60,
    MAPless55_5mins=sum(MAP <55, na.rm = TRUE)/60,
    MAPge55_64_5mins=sum(MAP >= 55 & MAP<64, na.rm = TRUE)/60,preredmins= sum((state=="R") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postredmins= sum((state=="R") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intraredmins= sum((state=="R") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_redmins= sum((state=="R") * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    preintra_red5mins= sum((state=="R") * (wftime <= cpboff_last), na.rm = TRUE) / 60,
    bluemins= sum(state=="B", na.rm = TRUE)  / 12,
    prebluemins= sum((state=="B") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postbluemins= sum((state=="B") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrabluemins= sum((state=="B") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_bluemins= sum((state=="B") * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    whitemins= sum(state=="W", na.rm = TRUE)  / 12,
    prewhitemins= sum((state=="W") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postwhitemins= sum((state=="W") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrawhitemins= sum((state=="W") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_whitemins= sum((state=="W") * (wftime <= cpboff_last), na.rm = TRUE) / 12, zone5mins= sum(zone5=="zone5", na.rm = TRUE)  / 12)%>% mutate(
    blue5mins_nobypass=(prebluemins+postbluemins)/5,
    white5mins_nobypass=(prewhitemins+postwhitemins)/5,
    red5mins_nobypass=(preredmins+postredmins)/5,
    intrablue5mins= intrabluemins/5,
    intrawhite5mins=intrawhitemins/5,
    intrared5mins=intraredmins/5) %>% mutate(timepoint=rep("OffCPB",n()))

icu_hemo =  hemodynamics %>% group_by(harbor_id) %>% mutate(harbor_id=as.integer(harbor_id)) %>% summarise(redzone5mins=sum(state=="R", na.rm = TRUE)  / 60,whitezone5mins=sum(state=="W", na.rm = TRUE)  / 60,bluezone5mins=sum(state=="B", na.rm = TRUE)  / 60, MAPless65_5mins= sum(MAP<65,na.rm=TRUE)/60,MAPge65_5mins=sum(MAP >= 65, na.rm = TRUE)/60,
    MAPless55_5mins=sum(MAP <55, na.rm = TRUE)/60,
    MAPge55_64_5mins=sum(MAP >= 55 & MAP<64, na.rm = TRUE)/60,preredmins= sum((state=="R") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postredmins= sum((state=="R") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intraredmins= sum((state=="R") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_redmins= sum((state=="R") * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    preintra_red5mins= sum((state=="R") * (wftime <= cpboff_last), na.rm = TRUE) / 60,
    bluemins= sum(state=="B", na.rm = TRUE)  / 12,
    prebluemins= sum((state=="B") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postbluemins= sum((state=="B") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrabluemins= sum((state=="B") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_bluemins= sum((state=="B") * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    whitemins= sum(state=="W", na.rm = TRUE)  / 12,
    prewhitemins= sum((state=="W") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postwhitemins= sum((state=="W") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrawhitemins= sum((state=="W") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_whitemins= sum((state=="W") * (wftime <= cpboff_last), na.rm = TRUE) / 12, zone5mins= sum(zone5=="zone5", na.rm = TRUE)  / 12)%>% mutate(
    blue5mins_nobypass=(prebluemins+postbluemins)/5,
    white5mins_nobypass=(prewhitemins+postwhitemins)/5,
    red5mins_nobypass=(preredmins+postredmins)/5,
    intrablue5mins= intrabluemins/5,
    intrawhite5mins=intrawhitemins/5,
    intrared5mins=intraredmins/5) %>% mutate(timepoint=rep("ICU",n())) 


biomarkers2= biomarkers2 %>% left_join(rbind(offcpb_hemo,icu_hemo),by=c("harbor_id","timepoint")) %>% drop_na() %>% left_join(biowide%>% select(harbor_id, cpbtime,proctime,predmort,predrenf),by="harbor_id")

biomarkers2=biomarkers2 %>% mutate(ngal_diff=log2NGAL_UCR-baseline_value)

```

    




```{r,echo=FALSE}

m1 <- lme(ngal_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort, random=~1 | harbor_id, data = biomarkers2)
tidy(m1, effects = "fixed", conf.int = TRUE, conf.level = 0.95)%>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))
```




```{r, include=FALSE, evaluate=FALSE}
m1 <- lme(ngal_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort, random=~1 | harbor_id, data = biomarkers2)
vif(m1)
```




```{r,echo=FALSE}

m1 <- lme(ngal_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort+MAPless65_5mins, random=~1 | harbor_id, data = biomarkers2)
tidy(m1, effects = "fixed", conf.int = TRUE, conf.level = 0.95)%>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))
```



```{r, include=FALSE, evaluate=FALSE}
m1 <- lme(ngal_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort+MAPless65_5mins, random=~1 | harbor_id, data = biomarkers2)
vif(m1)
```

```{r,echo=FALSE}


m1 <- lme(ngal_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort+cpbtime, random=~1 | harbor_id, data = biomarkers2)
tidy(m1, effects = "fixed", conf.int = TRUE, conf.level = 0.95)%>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))
```


```{r, include=FALSE, evaluate=FALSE}
m1 <- lme(ngal_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort+cpbtime, random=~1 | harbor_id, data = biomarkers2)
vif(m1)
```



### Mixed effects nephrocheck through 24 hours

```{r,echo=FALSE}
biomarkers2= biomarkers %>% filter(!timepoint  %in% c("PreCPB")) %>% select(harbor_id,timepoint,akiscore)
timepoint_1_values <- biomarkers2 %>%
  filter(timepoint == "Baseline") %>%
  select(harbor_id, baseline_value = akiscore)
biomarkers2 = biomarkers2%>%
  filter(timepoint != "Baseline") %>%
  left_join(timepoint_1_values, by = "harbor_id")

offcpb_hemo = hemodynamics %>% group_by(harbor_id) %>% filter(wftime<=cpboff_last) %>% mutate(harbor_id=as.integer(harbor_id))%>% summarise(redzone5mins=sum(state=="R", na.rm = TRUE)  / 60,whitezone5mins=sum(state=="W", na.rm = TRUE)  / 60,bluezone5mins=sum(state=="B", na.rm = TRUE)  / 60, MAPless65_5mins= sum(MAP<65,na.rm=TRUE)/60, MAPge65_5mins=sum(MAP >= 65, na.rm = TRUE)/60,
    MAPless55_5mins=sum(MAP <55, na.rm = TRUE)/60,
    MAPge55_64_5mins=sum(MAP >= 55 & MAP<64, na.rm = TRUE)/60,preredmins= sum((state=="R") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postredmins= sum((state=="R") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intraredmins= sum((state=="R") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_redmins= sum((state=="R") * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    preintra_red5mins= sum((state=="R") * (wftime <= cpboff_last), na.rm = TRUE) / 60,
    bluemins= sum(state=="B", na.rm = TRUE)  / 12,
    prebluemins= sum((state=="B") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postbluemins= sum((state=="B") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrabluemins= sum((state=="B") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_bluemins= sum((state=="B") * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    whitemins= sum(state=="W", na.rm = TRUE)  / 12,
    prewhitemins= sum((state=="W") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postwhitemins= sum((state=="W") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrawhitemins= sum((state=="W") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_whitemins= sum((state=="W") * (wftime <= cpboff_last), na.rm = TRUE) / 12, zone5mins= sum(zone5=="zone5", na.rm = TRUE)  / 12)%>% mutate(
    blue5mins_nobypass=(prebluemins+postbluemins)/5,
    white5mins_nobypass=(prewhitemins+postwhitemins)/5,
    red5mins_nobypass=(preredmins+postredmins)/5,
    intrablue5mins= intrabluemins/5,
    intrawhite5mins=intrawhitemins/5,
    intrared5mins=intraredmins/5)%>% mutate(timepoint=rep("OffCPB",n()))

icu_hemo =  hemodynamics %>% group_by(harbor_id) %>% mutate(harbor_id=as.integer(harbor_id)) %>% summarise(redzone5mins=sum(state=="R", na.rm = TRUE)  / 60,whitezone5mins=sum(state=="W", na.rm = TRUE)  / 60,bluezone5mins=sum(state=="B", na.rm = TRUE)  / 60, MAPless65_5mins= sum(MAP<65,na.rm=TRUE)/60, MAPge65_5mins=sum(MAP >= 65, na.rm = TRUE)/60,
    MAPless55_5mins=sum(MAP <55, na.rm = TRUE)/60,
    MAPge55_64_5mins=sum(MAP >= 55 & MAP<64, na.rm = TRUE)/60,preredmins= sum((state=="R") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postredmins= sum((state=="R") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intraredmins= sum((state=="R") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_redmins= sum((state=="R") * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    preintra_red5mins= sum((state=="R") * (wftime <= cpboff_last), na.rm = TRUE) / 60,
    bluemins= sum(state=="B", na.rm = TRUE)  / 12,
    prebluemins= sum((state=="B") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postbluemins= sum((state=="B") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrabluemins= sum((state=="B") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_bluemins= sum((state=="B") * (wftime <= cpboff_last), na.rm = TRUE) / 12,
    whitemins= sum(state=="W", na.rm = TRUE)  / 12,
    prewhitemins= sum((state=="W") * (wftime < cpb1on), na.rm = TRUE) / 12,
    postwhitemins= sum((state=="W") * (wftime > cpboff_last), na.rm = TRUE) / 12,
    intrawhitemins= sum((state=="W") * (wftime < cpboff_last & wftime>cpb1on), na.rm = TRUE) / 12,
    preintra_whitemins= sum((state=="W") * (wftime <= cpboff_last), na.rm = TRUE) / 12, zone5mins= sum(zone5=="zone5", na.rm = TRUE)  / 12)%>% mutate(
    blue5mins_nobypass=(prebluemins+postbluemins)/5,
    white5mins_nobypass=(prewhitemins+postwhitemins)/5,
    red5mins_nobypass=(preredmins+postredmins)/5,
    intrablue5mins= intrabluemins/5,
    intrawhite5mins=intrawhitemins/5,
    intrared5mins=intraredmins/5)%>% mutate(timepoint=rep("ICU",n()))

icu24h_hemo =  icu_hemo %>% mutate(timepoint=rep("24hICU",n()))

biomarkers2= biomarkers2 %>% left_join(rbind(offcpb_hemo,icu_hemo,icu24h_hemo),by=c("harbor_id","timepoint")) %>% drop_na() %>% left_join(biowide%>% select(harbor_id, cpbtime,proctime,predmort,predrenf),by="harbor_id")

biomarkers2=biomarkers2 %>% mutate(nephro_diff=akiscore-baseline_value)

```






```{r,echo=FALSE}

m1 <- lme(nephro_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort, random=~1 | harbor_id, data = biomarkers2)
tidy(m1, effects = "fixed", conf.int = TRUE, conf.level = 0.95)%>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))
```


```{r, include=FALSE, evaluate=FALSE}
m1 <- lme(nephro_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort, random=~1 | harbor_id, data = biomarkers2)
vif(m1)
```


```{r,echo=FALSE}

m1 <- lme(nephro_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort+MAPless65_5mins, random=~1 | harbor_id, data = biomarkers2)
tidy(m1, effects = "fixed", conf.int = TRUE, conf.level = 0.95)%>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))
```


```{r, include=FALSE, evaluate=FALSE}
m1 <- lme(nephro_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort+MAPless65_5mins, random=~1 | harbor_id, data = biomarkers2)
vif(m1)
```

```{r,echo=FALSE}

m1 <- lme(nephro_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort+cpbtime, random=~1 | harbor_id, data = biomarkers2)
tidy(m1, effects = "fixed", conf.int = TRUE, conf.level = 0.95)%>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))
```

```{r, include=FALSE, evaluate=FALSE}
m1 <- lme(nephro_diff ~ redzone5mins +whitezone5mins+bluezone5mins+predmort+cpbtime, random=~1 | harbor_id, data = biomarkers2)
vif(m1)
```






```{r,echo=FALSE}
cor_matrix <- biowide %>%
  select(predmort, red5mins, blue5mins, white5mins, MAPless65_5mins, cpbtime) %>%
  cor(use = "pairwise.complete.obs") 

ggcorrplot(cor_matrix, 
           type = "lower",      # Only show the bottom half
           lab = TRUE,          # Show the actual numbers
           title = "")

```

## Mixed Effects MAP<65 and CVP>12 (baseline to ICU arrival)


```{r,echo=FALSE}
summary_df = hemodynamics %>%
  mutate(
    map65 = MAP < 65,
    cvp12=CVP>12
  ) %>%
  group_by(harbor_id) %>%
  summarise(
    map65_5minutes = sum(map65, na.rm = TRUE) / 60,
    cvp12_5minutes = sum(cvp12, na.rm = TRUE) / 60,
  )
biowide= biowide %>% left_join(summary_df, by="harbor_id")

```

```{r,echo=FALSE}
biomarkers2= biomarkers %>% filter(!timepoint  %in% c("PreCPB","24hICU","OffCPB")) %>% select(harbor_id,timepoint,log2NGAL_UCR)
timepoint_1_values <- biomarkers2 %>%
  filter(timepoint == "Baseline") %>%
  select(harbor_id, baseline_value = log2NGAL_UCR)
biomarkers2 = biomarkers2%>%
  filter(timepoint != "Baseline") %>%
  left_join(timepoint_1_values, by = "harbor_id")


biomarkers2= biomarkers2 %>% left_join(biowide%>% select(harbor_id, cpbtime,proctime, map65_5minutes, cvp12_5minutes),by="harbor_id")

biomarkers2=biomarkers2 %>% mutate(ngal_diff=log2NGAL_UCR-baseline_value)
```


log2ngal_ucr at icu arrival - baseline

```{r,echo=FALSE}
m1 <- lme(ngal_diff ~ map65_5minutes  , random=~1 | harbor_id, data = biomarkers2)
tidy(m1, effects = "fixed", conf.int = TRUE, conf.level = 0.95)%>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))

m1 <- lme(ngal_diff ~ cvp12_5minutes  , random=~1 | harbor_id, data = biomarkers2)
tidy(m1, effects = "fixed", conf.int = TRUE, conf.level = 0.95)%>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))


m1 <- lme(ngal_diff ~ map65_5minutes+cvp12_5minutes  , random=~1 | harbor_id, data = biomarkers2)
tidy(m1, effects = "fixed", conf.int = TRUE, conf.level = 0.95)%>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))
```




## Table 1: Log2NGAL_UCR tertiles at ICU Arrival 

```{r,echo=FALSE}
biowide <- biowide %>%
  mutate(log2NGAL_UCR_2_tertiles = ntile(log2NGAL_UCR_2, 3))

quantile(biowide$log2NGAL_UCR_2, probs = seq(0, 1, by = 1/3), na.rm = TRUE)
```



1 = lowest tertile = below -0.733  
3 = highest tertile = above 3.2



```{r,echo=FALSE}
covs=biowide %>% dplyr::select(log2NGAL_UCR_2_tertiles,patientage,gender,bmi,rfhypertension,rfdiabetes,priormi,hemodataef,rfpriorcva ,rfchroniclungdisease,medsbetablockerswithin24hours, rfhemoglobin,rflastcreatlevel,proctime,cpbtime,packed_rbc, predmort, predrenf, heartfailure, vasopressor, total_crystalloid,medslipidloweringwithin24hours,medsaceinhibitorsorarbwithin48ho)

covs %>%
  tbl_summary(
    by = log2NGAL_UCR_2_tertiles,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p()

covs %>%
  tbl_summary(
    by = log2NGAL_UCR_2_tertiles,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p() %>%
  as_tibble() %>%
  write.csv("log2ngalucrtertiles_table.csv", row.names = FALSE)

rm(covs)


```

```{r,echo=FALSE}

# Test normality of the continuous variables 
# None are normal if p-value<0.05
covs=biowide %>% dplyr::select(log2NGAL_UCR_2_tertiles,patientage, bmi, hemodataef, rfhemoglobin,rflastcreatlevel,proctime,cpbtime,predmort, predrenf,vasopressor, total_crystalloid)
shapiro_results <- lapply(covs, function(x) {
  x <- na.omit(x)   # remove missing values
  shapiro.test(x)
})

# Extract p-values
sapply(shapiro_results, function(x) x$p.value)
```



## Table 1: By Nephrocheck Thresholds

```{r}
covs=biowide %>% dplyr::select(akiscore_3_thresh03,patientage,gender,bmi,rfhypertension,rfdiabetes,priormi,hemodataef,rfpriorcva ,rfchroniclungdisease,medsbetablockerswithin24hours, rfhemoglobin,rflastcreatlevel,proctime,cpbtime,packed_rbc, predmort, predrenf, heartfailure, vasopressor, total_crystalloid,medslipidloweringwithin24hours,medsaceinhibitorsorarbwithin48ho)


covs %>%
  tbl_summary(
    by = akiscore_3_thresh03,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p()


covs=biowide %>% dplyr::select(akiscore_3_thresh2,patientage,gender,bmi,rfhypertension,rfdiabetes,priormi,hemodataef,rfpriorcva ,rfchroniclungdisease,medsbetablockerswithin24hours, rfhemoglobin,rflastcreatlevel,proctime,cpbtime,packed_rbc, predmort, predrenf, heartfailure, vasopressor, total_crystalloid,medslipidloweringwithin24hours,medsaceinhibitorsorarbwithin48ho)


covs %>%
  tbl_summary(
    by = akiscore_3_thresh2,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p()

```




## Table 1: Nephrocheck tertiles at ICU arrival 


```{r,echo=FALSE}
biowide <- biowide %>%
  mutate(akiscore_2_tertiles = ntile(akiscore_2, 3))

quantile(biowide$akiscore_2, probs = seq(0, 1, by = 1/3), na.rm = TRUE)
```



1 = lowest tertile = below 0.06
3 = highest tertile = above 0.116


```{r,echo=FALSE}
covs=biowide %>% dplyr::select(akiscore_2_tertiles,patientage,gender,bmi,rfhypertension,rfdiabetes,priormi,hemodataef,rfpriorcva ,rfchroniclungdisease,medsbetablockerswithin24hours, rfhemoglobin,rflastcreatlevel,proctime,cpbtime,packed_rbc, predmort, predrenf, heartfailure, vasopressor, total_crystalloid,medslipidloweringwithin24hours,medsaceinhibitorsorarbwithin48ho)

covs %>%
  tbl_summary(
    by = akiscore_2_tertiles,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p()

covs %>%
  tbl_summary(
    by = akiscore_2_tertiles,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p()%>%
  as_tibble() %>%
  write.csv("nephrocheck_tertiles_icuarrival_table.csv", row.names = FALSE)

rm(covs)

```


## Table 1: Nephrocheck tertiles at 24hICU


```{r,echo=FALSE}
biowide <- biowide %>%
  mutate(akiscore_3_tertiles = ntile(akiscore_3, 3))

quantile(biowide$akiscore_3, probs = seq(0, 1, by = 1/3), na.rm = TRUE)
```



1 = lowest tertile = below 0.5
3 = highest tertile = above 1.193


```{r,echo=FALSE}
covs=biowide %>% dplyr::select(akiscore_3_tertiles,patientage,gender,bmi,rfhypertension,rfdiabetes,priormi,hemodataef,rfpriorcva ,rfchroniclungdisease,medsbetablockerswithin24hours, rfhemoglobin,rflastcreatlevel,proctime,cpbtime,packed_rbc, predmort, predrenf, heartfailure, vasopressor, total_crystalloid,medslipidloweringwithin24hours,medsaceinhibitorsorarbwithin48ho)

covs %>%
  tbl_summary(
    by = akiscore_3_tertiles,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p()

covs %>%
  tbl_summary(
    by = akiscore_3_tertiles,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p()%>%
  as_tibble() %>%
  write.csv("nephrocheck_tertiles_icu24h_table.csv", row.names = FALSE)

rm(covs)

```

## Table 1: Nephrocheck at 24 h vs those who have missing


```{r,echo=FALSE}

biowide <- biowide %>%
  mutate(nephro24_missing = ifelse(is.na(akiscore_3), 1, 0))

covs=biowide %>% dplyr::select(nephro24_missing,patientage,gender,bmi,rfhypertension,rfdiabetes,priormi,hemodataef,rfpriorcva ,rfchroniclungdisease,medsbetablockerswithin24hours, rfhemoglobin,rflastcreatlevel,proctime,cpbtime,packed_rbc, predmort, predrenf, heartfailure, vasopressor, total_crystalloid,medslipidloweringwithin24hours,medsaceinhibitorsorarbwithin48ho,redmins )

covs %>%
  tbl_summary(
    by = nephro24_missing,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p()

```






## Table 1: Total Surgery Zone 3 tertiles 


```{r,echo=FALSE}
biowide <- biowide %>%
  mutate(whitemins_tertiles = ntile(whitemins, 3))

quantile(biowide$whitemins, probs = seq(0, 1, by = 1/3), na.rm = TRUE)
```

1 = lowest tertile = below 131.58333
3 = highest tertile = above 179.47222


```{r,echo=FALSE}
covs=biowide %>% dplyr::select(whitemins_tertiles,patientage,gender,bmi,rfhypertension,rfdiabetes,priormi,hemodataef,rfpriorcva ,rfchroniclungdisease,medsbetablockerswithin24hours, rfhemoglobin,rflastcreatlevel,proctime,cpbtime,packed_rbc, predmort, predrenf, heartfailure, vasopressor, total_crystalloid,medslipidloweringwithin24hours,medsaceinhibitorsorarbwithin48ho)

covs %>%
  tbl_summary(
    by = whitemins_tertiles,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p()

covs %>%
  tbl_summary(
    by = whitemins_tertiles,
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing_text = "Missing"
  ) %>%
  add_p()%>%
  as_tibble() %>%
  write.csv("zone3whitemins_tertiles_table.csv", row.names = FALSE)

rm(covs)

```



# Miscellaneous Revisions  

1)	Proportion of patients >0.3 and >2.0 raw nephrocheck at 
a.	End of surgery
b.	At 24 hours

```{r}
sum(biowide$akiscore_2>0.3)/nrow(biowide)
sum(biowide$akiscore_2>2)/nrow(biowide)

sum(biowide$akiscore_3>0.3,na.rm=TRUE)
sum(biowide$akiscore_3>2,na.rm=TRUE)
nrow(biowide)- sum(is.na(biowide$akiscore_3))
```



```{r}
sum(biowide$akiscore_2>0.3 & biowide$AKI_48==1)
sum(biowide$akiscore_2>2 & biowide$AKI_48==1)


sum(biowide$akiscore_3>0.3& biowide$AKI_48==1,na.rm=TRUE)
sum(biowide$akiscore_3>2& biowide$AKI_48==1,na.rm=TRUE)
```





2)	Please explore (as requested by the authors): relationship between total time in zone 3 and (yes/no above vs below threshold) logistic regression. Please adjust for STS, include zone 2 and zone 1. (please do for both thresholds at both timepoints
e
```{r}


tidy(
  glm(
    akiscore_3_thresh03 ~ red5mins + white5mins + blue5mins + predmort,
    data = biowide,
    family = binomial
  ),
  exponentiate = TRUE,
  conf.int = TRUE
) %>%
  mutate(
    across(c(estimate, conf.low, conf.high), ~ round(.x, 3)),
    p.value = round(p.value, 3)
  )



tidy(
  glm(
    akiscore_3_thresh2 ~ red5mins + white5mins + blue5mins + predmort,
    data = biowide,
    family = binomial
  ),
  exponentiate = TRUE,
  conf.int = TRUE
) %>%
  mutate(
    across(c(estimate, conf.low, conf.high), ~ round(.x, 3)),
    p.value = round(p.value, 3)
  )
```






```{r}

tidy(lm(akiscore_2 ~ red5mins + white5mins + blue5mins + predmort, data=biowide))
tidy(lm(akiscore_2 ~ red5mins + white5mins + blue5mins + predmort + cpbtime, data=biowide))


tidy(lm(akiscore_3 ~ red5mins + white5mins + blue5mins + predmort, data=biowide))
tidy(lm(akiscore_3 ~ red5mins + white5mins + blue5mins + predmort + cpbtime, data=biowide))
```





```{r}


tidy(glm(akiscore_2_thresh03 ~ red5mins+white5mins+blue5mins+predmort, data = biowide, family = "binomial"))
tidy(glm(akiscore_2_thresh2 ~ red5mins+white5mins+blue5mins+predmort, data = biowide, family = "binomial"))
```




```{r}
hemodynamics %>%
  filter(state == "R") %>%
  mutate(cat_cpb = fct_recode(cat_cpb,
    "Pre-Bypass" = "pre",
    "Intra-Bypass" = "intra",
    "Post-Bypass" = "post"
  )) %>%
  group_by(cat_cpb, harbor_id) %>%
  summarize(total_minutes = n() / 12) %>% 
  ggplot(aes(x=cat_cpb,y=total_minutes))+geom_boxplot()+ylab("Total Minutes in Zone 3")+xlab("Surgical Period")+theme_minimal()
```



